\hypertarget{index_intro}{}\section{Introduction}\label{index_intro}
{\bf GenomicTools} is a flexible computational platform for the analysis and manipulation of high-throughput sequencing data such as DNA-seq, RNA-seq, ChIP-seq and MethylC-seq. GenomicTools implements a variety of mathematical operations between sets of genomic regions thereby enabling the prototyping of computational pipelines that can address a wide spectrum of tasks from preprocessing and quality control to meta-analyses. For example, the user can easily create average read profiles across transcriptional start sites or enhancer sites, quickly prototype customized peak discovery methods for ChIP-seq experiments, perform genome-wide statistical tests such as enrichment analyses, design controls via appropriate randomization schemes, among other applications. In addition to enabling rapid prototyping, the GenomicTools platform is designed to analyze large-datasets of any size by minimizing memory requirements. Compared to Bioconductor packages, it offers a $\sim$10-fold improvement by relying on single-pass processing and fast I/O implementation. When analyzing hundreds of patient datasets, this is a significant advantage. Finally, the GenomicTools platform supports the widely used BED format to facilitate visualization as well as integration with existing platforms and pipelines such as Galaxy or Bioconductor.\hypertarget{index_download}{}\section{Download and install}\label{index_download}
{\bf GenomicTools} can be downloaded from \href{http://http://code.google.com/p/ibm-cbc-genomic-tools/}{\tt code.google.com}. To install it, simply enter the src subdirectory and type make. This will create the executables in the bin subdirectory (which you may choose to included in your path). 